---
title: 'Tidy Tuesday Homework 2: Week of Nov. 5th'
author: "Diana Gao"
date: "2024-11-12"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../output/") #set directory for the figures to go in to
```

# Setup

#### Load the Libraries 
```{r}
library(tidytuesdayR)
library(tidyverse)
library(here)
library(gganimate)
```


#### Load in Data
This week I will be using the "Democracy and Dictatorship" dataset from TidyTuesday, week of Nov. 5th
```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 45)
democracy_data <- tuesdata$democracy_data
```

# Cleaning the Data 
My idea is to plot the number of democratic, communist, and monarchy governments every year and to animate this with gganimate.

Today the goal is to learn to use gganimate with line geoms. I struggled to correctly count the number of is_xxx = TRUE as I was trying the count() function but it said it didn't work on logical data. 
```{r}
clean <- democracy_data %>%
  select(country_name, year, is_democracy, is_communist, is_monarchy) %>% # keep only desired cols
  group_by(year) %>%
  mutate(democracy = sum(is_democracy, na.rm = TRUE), 
            communist = sum(is_communist, na.rm = TRUE), 
            monarchy = sum(is_monarchy, na.rm = TRUE)) %>% # count how many of each regime there are per year
  select(year, democracy, communist, monarchy) %>% # only keep the relevant columns
  pivot_longer(!year, names_to = "regime", values_to = "count") # pivot data to longer so they can all be plotted at once with a line geom
```

# Plotting
I also struggled a bit to combine plots of all three different regimes before realizing I had to make the data into a longer format. The plotting went smoothly once I figured out the correct animation function (transition_reveal)
```{r, regimes_over_time}
animation <- clean %>%
  ggplot(aes(x = year, y = count, 
             color = regime)) +
  geom_line() + 
  labs(title = "Number of democratic, communist, and monarchist countries over time", 
       caption = "Data is from tidyTuesday 2024-11-15", 
       x = "Year", 
       y = "Number of Nations") + 
  theme_bw() +
  transition_reveal(year)

animate(animation, nframes = 200, #make into 200 frames
        fps = 24, #set the framerate to be 24 frames per seconds
)
```

