---
title: "Week 7 At Home Lecture Notes"
author: "Diana Gao"
date: "2024-10-15"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../output/") #set directory for the figures to go in to
```

# Setup

#### Load Libraries

```{r}
library(tidyverse)
library(hms)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(gganimate)
library(gifski)
library(tidytuesdayR)
#library(CatterPlots) # I really wanted to make the points a cat moving around but couldnt figure it out lol
```

#### Read in the Data
I chose to use the data from 2023-01-31, "cats on the move!" This data tracks the movement of cats in the UK :)

```{r}
tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference
```
#### Filter the Data
I decided to follow coco and map her journey :)
```{r}
is.character(cats_uk$timestamp) #character is no, must change to character
cats_uk$timestamp <- as.character(cats_uk$timestamp) # change timestamp to character
#is.character(cats_uk$timestamp) #Just to check again :)
coco_data <- cats_uk %>%
  mutate(time = ymd_hms(timestamp)) %>% # for some reason wasn't working when I named column timestamp?? so I just made new column named time
  mutate(only_time = as_hms(time)) %>% # make a col with only the time
  mutate(day = round_date(time, "day")) %>% # make a col with only the date
  filter(tag_id == "Coco") # select only coco's data points

```

# Plotting

#### Make map of UK
Here is where Coco lives within the UK 
```{r, where_is_coco}
UK <- map_data(map = "world", region = "UK") # select the UK region out of the world map, unfortunately subregion is only based on landmasses ^^'
ggplot() + 
  geom_polygon(data = UK, aes(x = long, y = lat, group = group)) +
  geom_point(data = coco_data, aes(x = location_long, y = location_lat), 
             color = "red", size = 1) +
  coord_map()
```

#### Plotting Coco's movement
Looking at coco_data, can see that her movement is limited to x = -5.077785, -5.073802 and y = 50.15111, 50.15638. So I will plot UK and zoom in to that general area so we can have enough spatial resolution to animate her moving around
```{r}
coco_moving <- ggplot() + 
  geom_point(data = coco_data, aes(x = location_long, y = location_lat), 
             color = "red", size = 3) +
  lims (x = c(-5.0785, -5.072), y = c(50.15, 50.157)) +
  coord_map() +
  facet_wrap(~day) + # separate graphs for each day
  labs(title = 'Hour: {frame_time}', x = '', y = '') + #remove x and y axes and set title for each
  transition_time(only_time) + # animate by the data in time column
  ease_aes('linear') # 
```
#### Render the animation
```{r, cocomoving}
animate(coco_moving, nframes = 200, #make into 200frames
        fps = 24, #set the frame by 20 frame per seconds
        width = 10, height = 4, units = "in",  #set the width and length of the frames using px
        renderer = gifski_renderer()) # render w gifski

```

