---
title: 'Tidy Tuesday Homework 4: Week of Nov. 26th'
author: "Diana Gao"
date: "2024-12-1"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../output/") #set directory for the figures to go in to
```

As introduced in Dr. Silbiger's slack, I will be trying out tidyplots today! 

# Setup
### Load the libraries 
```{r}
library(tidyverse)
library(tidyplots)
library(here)
```

### Load in the data 
```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 48)

cbp_resp <- tuesdata$cbp_resp
cbp_state <- tuesdata$cbp_state
```

### Data cleaning

I will be looking at the number of individuals encountered at all borders, looking to see if the amount changed from 2020-2024 and if the title they were processed under changed. Thus, I only need a few columns of data and the overall number of individuals per year per title (rather than number of individuals for each encounter)

```{r}
cbp_clean <- cbp_resp %>%
  select(fiscal_year, title_of_authority, encounter_count) %>% # only save these cols
  group_by(title_of_authority, fiscal_year) %>% # group by both title of authority and fiscal year
  count() # count the number of individuals by the groups above

```

# Plotting

Now for the new stuff, trying out tidyplots! I made a stacked bar chart: 
```{r, title_count_per_year.png}
cbp_clean %>%
  tidyplot(x = fiscal_year, 
           y = n, 
           color = title_of_authority) %>%
  add_barstack_absolute() %>% # stacked barchart geom
  adjust_title("Number of individuals encountered per fiscal year") %>% # changing names and stuff
  adjust_x_axis_title("Fiscal Year") %>%
  adjust_y_axis_title("Number of Individuals Encountered") %>%
  adjust_legend_title("Title of Authority") #%>%
  # save_plot(here("tidy_tuesday", "5_Nov26", "output", "titles.png")) # in basic code I would save using this function :)
```

The "structure" of the code is definetly different from ggplot, but I really like that you can use the maggittr pipe with it! Also, I feel like the commands are more intuitively named, and thus easier to search and find. Overall, I like it! 